---
title: "Predict with Model & Geostats"
output: html_document
---

This notebook combines the statistical and geostatistical models to make a 
prediction of geomechanical properties across the basin. Steps include:

- Load stress realizations and log inputs realizations (carbonate + shale)
- Estimate required stresses (confining + deviatoric) using stress polygon method
- Load a model
- Make predictions on each realization

```{r setup}
library(duvernaygeomechmodel)
library(stars)
study_bounds = st_read('../data/study_dv_bounds.GeoJSON')
study_raster = raster::raster('../data/study_raster.tif')
```

Rasterize sv_sf for mbgs datum (needed for absolute stress conversion)

```{r}
sv_sf = st_read('../data/sv_sf.GeoJSON') %>% 
  drop_na() %>% 
  select(datum_mbgs)

datum_mbgs = st_rasterize(sv_sf, template = st_as_stars(study_raster)) %>% 
  as(., 'Raster')
```

Load stress realizations and log inputs realizations (carbonate + shale)

```{r}
stress = readRDS('../output/geostatistical_models/stress_pred_trans.rds')
shale_logs = readRDS('../output/geostatistical_models/shale_logs_pred_trans.rds')
carb_logs = readRDS('../output/geostatistical_models/carb_logs_pred_trans.rds')

# make a predictions spdf to contain predictions
predictions = stress
predictions@data$x = predictions@coords[,1]
predictions@data$y = predictions@coords[,2]
predictions@data <- predictions@data %>% select(x,y)

```

Use Stress Polygon method to predict SHmax. 

The polygon method requires an R-factor,
which we can either set deterministically, or simulate it by drawing values from
a distribution. Since we lack data to constrain this distribution, we use a 
deterministic assumption of r = 0.62. This produces mean shmax values across the 
simulations of 33.3 +/- 1 kpa/m, which is consistent with the value reported by
Shen et al. (2019). Without further constraints, simulating the r-factor is
roughly equivalent to adding random noise into the workflow and of little 
interpretational benefit.

We also calculate deviatoric and confining stress, assuming a Biot coefficient
of 0.7. 

```{r}
r=0.62

sim_vec = str_replace_all(
  colnames(stress@data)[grepl("sv.*", colnames(stress@data))],
  "sv.",""
  )

for (sim in sim_vec){
  shmin = stress@data[, paste0('shmin.',sim)]
  sv = stress@data[, paste0('sv.',sim)]
  pp = stress@data[, paste0('pp.',sim)]
  sigma_2 = ifelse(sv > shmin, sv, shmin)
  sigma_3 = ifelse(sv > shmin, shmin, sv)
  # stress polygon calc
  sigma_1 = (sigma_3 * r - sigma_2)/(r - 1)
  stress@data[, paste0("shmax.", sim)] <- sigma_1
  # confining and deviatoric stress calcs
  stress@data[, paste0("sigc.", sim)] <- (sigma_2 + sigma_3)/2 - 0.7 * pp
  stress@data[, paste0("sigd.", sim)] <- sigma_1 - sigma_3
}
```

Check mean stress values for calculated parameters

```{r}
# find mean & sd of shmax values
shmax_summary = stress@data %>%
  rowwise() %>%
  summarize(
    shmax_mean = mean(c_across(contains('shmax'))),
    shmax_sd = sd(c_across(contains('shmax'))),
    sigc_mean = mean(c_across(contains('sigc.'))),
    sigc_sd = sd(c_across(contains('sigc.'))),
    sigd_mean = mean(c_across(contains('sigd.'))),
    sigd_sd = sd(c_across(contains('sigd.')))
  )

mean(shmax_summary$shmax_mean)
sd(shmax_summary$shmax_mean)
mean(shmax_summary$sigc_mean)
sd(shmax_summary$sigc_sd)
mean(shmax_summary$sigd_mean)
sd(shmax_summary$sigd_sd)
```
Convert stresses to absolute values

```{r}
# functions to make absolute stress
grad_to_abs_val <- function(x, datum_mbgs){
  x * datum_mbgs@data@values / 1000
}

stress_abs = stress
stress_abs@data <- stress@data %>%
  mutate(across(contains('sv'), grad_to_abs_val, datum_mbgs)) %>%
  mutate(across(contains('shmin'), grad_to_abs_val, datum_mbgs)) %>%
  mutate(across(contains('pp'), grad_to_abs_val, datum_mbgs)) %>%
  mutate(across(contains('shmax'), grad_to_abs_val, datum_mbgs)) %>%
  mutate(across(contains('sigc'), grad_to_abs_val, datum_mbgs)) %>%
  mutate(across(contains('sigd'), grad_to_abs_val, datum_mbgs)) %>%
  na_if(0)
```

Convert logs to vp and vs and remove NA areas

```{r}
# convert slowness to velocity and remove masked areas
dt_to_v <- function(x, datum_mbgs){
  valid_mask = ifelse(datum_mbgs@data@values>0,1,0)
  x^-1*1e6*valid_mask
}

# convert rhob to density and remove masked areas
rhob_to_density <- function(x, datum_mbgs){
  valid_mask = ifelse(datum_mbgs@data@values>0,1,0)
  x/1000*valid_mask
}

shale_logs_v = shale_logs
shale_logs_v@data <- shale_logs@data %>%
  mutate(across(contains('rhob'), rhob_to_density, datum_mbgs)) %>%
  mutate(across(contains('dtc'), dt_to_v, datum_mbgs)) %>%
  mutate(across(contains('dts'), dt_to_v, datum_mbgs)) %>%
  na_if(0)

colnames(shale_logs_v@data) <- colnames(shale_logs_v@data) %>%
  str_replace_all(., "dtc", "vp") %>%
  str_replace_all(., "dts", "vs") %>%
  str_replace_all(., "rhob", "rho")

carb_logs_v = carb_logs
carb_logs_v@data <- carb_logs@data %>%
  mutate(across(contains('rhob'), rhob_to_density, datum_mbgs)) %>%
  mutate(across(contains('dtc'), dt_to_v, datum_mbgs)) %>%
  mutate(across(contains('dts'), dt_to_v, datum_mbgs)) %>%
  na_if(0)

colnames(carb_logs_v@data) <- colnames(carb_logs_v@data) %>%
  str_replace_all(., "dtc", "vp") %>%
  str_replace_all(., "dts", "vs") %>%
  str_replace_all(., "rhob", "rho")
```

Plot individual realizations

```{r}
sp_col_to_df <- function(spdf, col){
  x = spdf@coords[,1]
  y = spdf@coords[,2]
  data = spdf@data[,col]
  df_out = data.frame(x,y,data)
  names(df_out) = c('x', 'y', col)
  df_out
}

col = 'sigc.sim2'
ggplot(data = sp_col_to_df(stress_abs, col)) +
  geom_tile(aes(x=x,y=y, fill=get(col)))
```
Iterate through each of the 1000 realizations, load the 54 models, and make 
predictions for shale and carbonate. This loop takes a long time to run...

We predict for vertical orientation only.

```{r}
for (sim in sim_vec){
  print(sim)
  
  # stresses
  deviatoric_stress = stress_abs@data[, paste0('sigd.',sim)]
  confining_pressure = stress_abs@data[, paste0('sigc.',sim)]
  
  # shale logs
  pred_vs = shale_logs_v@data[, paste0('vs.',sim)]
  pred_vp = shale_logs_v@data[, paste0('vp.',sim)]
  bulk_density = shale_logs_v@data[, paste0('rho.',sim)]
  
  na_mask = ifelse(is.na(pred_vs), NA, 1)
  shale_df_out = data.frame(
    deviatoric_stress , confining_pressure, pred_vs, pred_vp, bulk_density
    ) %>% replace(is.na(.), 0) %>%
    mutate(shale_group = 1) %>%
    mutate(horiz_orient= 0)

  # carbonate logs
  pred_vs = carb_logs_v@data[, paste0('vs.',sim)]
  pred_vp = carb_logs_v@data[, paste0('vp.',sim)]
  bulk_density = carb_logs_v@data[, paste0('rho.',sim)]
  
  carb_df_out = data.frame(
    deviatoric_stress , confining_pressure, pred_vs, pred_vp, bulk_density
    ) %>% replace(is.na(.), 0) %>%
    mutate(shale_group = 0) %>%
    mutate(horiz_orient= 0)
  
  # cycle through models and make predictions for shale and carbonate
  # load model(s)
  for (dir in list.dirs("../output/statistical_models/",recursive = FALSE)){
    model_name = tail(str_split(dir,'/')[[1]],1)
    model = readRDS(paste0(dir,'/resample_results.rds'))$model
    predictions@data[, paste0('shale.',model_name,".",sim)] <- predict(
      model, newdata=shale_df_out %>% select(model$features))$data$response * na_mask
    predictions@data[, paste0('carb.',model_name,".",sim)] <- predict(
      model, newdata=carb_df_out %>% select(model$features))$data$response * na_mask
  }
}

saveRDS(predictions,'../output/geostatistical_models/predictions.rds')
```
Plot some predictions - looks as expected, with lots of variance when the 
geostatistical variance is applied the models! This is good and will be very
interesting to write up.

```{r}
colnames(predictions@data)

col = 'carb.dev_conf_vp_rho_rf_br.sim3'

ggplot(data = sp_col_to_df(predictions, col)) +
  geom_tile(aes(x=x,y=y, fill=get(col)))

```

- Summarize predictions (mean + sd + CI90)

```{r}

```



