---
title: "Data Preparation"
output: html_document
---

```{r setup, include=FALSE}
library(duvernaygeomechmodel)
```

This notebook loads static (unconfined compressive strength and triaxial) and 
dynamic (ultrasonic pulse transmission) data. It merges these two datasets
in preparation for statistical model generation and does some exploratory data
analysis.

# Load files

```{r}
static_df = read_csv('../data/static_tests.csv')
upt_df = read_csv('../data/upt_tests.csv')
```

# Merge triaxial and ultrasonic data

First we nest the static data, omitting null values (mainly from bulk density).
and grouping by well, depth, orientation, confining pressure, and deviatoric 
stress. We fit a model of Vp and Vs against confining and deviatoric stress for each 
ultrasonic measurement.

```{r}
static_nest_no_res <- static_df %>% 
  dplyr::select(-yield_stress_dev, -residual_stress_dev) %>%
  mutate(deviatoric_stress = peak_stress_dev / 2) %>%
  na.omit() %>%
  mutate(conf_pres_group = confining_pressure, 
         dev_stress_group = deviatoric_stress,
         well_group = well,
         depth_group = round_depth,
         orient_group = orientation) %>%
  group_by(well_group, depth_group, orient_group, conf_pres_group, dev_stress_group) %>%
  nest() %>%
  rename(confining_pressure = conf_pres_group, 
         deviatoric_stress = dev_stress_group,
         well = well_group,
         round_depth = depth_group,
         orientation = orient_group)

static_nest_w_res <- static_df %>% 
  dplyr::select(-yield_stress_dev) %>%
  mutate(deviatoric_stress = peak_stress_dev / 2) %>%
  na.omit() %>%
  mutate(conf_pres_group = confining_pressure, 
         dev_stress_group = deviatoric_stress,
         well_group = well,
         depth_group = round_depth,
         orient_group = orientation) %>%
  group_by(well_group, depth_group, orient_group, conf_pres_group, dev_stress_group) %>%
  nest() %>%
  rename(confining_pressure = conf_pres_group, 
         deviatoric_stress = dev_stress_group,
         well = well_group,
         round_depth = depth_group,
         orientation = orient_group)

upt_nest <- upt_df %>%
  group_by(well, round_depth, orientation) %>%
  nest() %>%
  mutate(vp_model = map(data, fit_vp_model)) %>%
  mutate(vs_model = map(data, fit_vs_model))
```

We then use these models to provide a prediction of Vp and Vs for the conditions
present in each triaxial test, for the static dataframes with and without residual
stresses. This chunk merges the tests with or without residual stress results.

```{r}
vp_df <- upt_nest %>%
  inner_join(static_nest_no_res, by = c('well','round_depth','orientation')) %>%
  rename(upt_data = data.x, trx_data = data.y) %>%
  map2_df(.x = .$vp_model, .y = .$trx_data, 
          .f = ~augment(.x, newdata = .y, se_fit=TRUE)) %>%
  rename(pred_vp = .fitted, se_vp = .se.fit)

vs_df <- upt_nest %>%
  inner_join(static_nest_no_res, by = c('well','round_depth','orientation')) %>%
  rename(upt_data = data.x, trx_data = data.y) %>%
  map2_df(.x = .$vs_model, .y = .$trx_data, 
          .f = ~augment(.x, newdata = .y, se_fit=TRUE)) %>%
  rename(pred_vs = .fitted, se_vs = .se.fit) %>%
  dplyr::select(pred_vs, se_vs)

merged_no_res <-  cbind(vp_df, vs_df) %>%
  mutate(se_vp_per = se_vp/pred_vp) %>%
  mutate(se_vs_per = se_vs/pred_vs) %>%
  filter(pred_vp >0, pred_vs >0) %>%
  filter(!is.na(se_vp), !is.na(se_vs)) %>%
  filter(se_vp_per < 0.1) %>%
  filter(se_vs_per < 0.1)

write_csv(merged_no_res, '../data/merged_tests_no_residual.csv')
```

This chunk merges the tests with residual stress results only.

```{r}
vp_df <- upt_nest %>%
  inner_join(static_nest_w_res, by = c('well','round_depth','orientation')) %>%
  rename(upt_data = data.x, trx_data = data.y) %>%
  map2_df(.x = .$vp_model, .y = .$trx_data, 
          .f = ~augment(.x, newdata = .y, se_fit=TRUE)) %>%
  rename(pred_vp = .fitted, se_vp = .se.fit)

vs_df <- upt_nest %>%
  inner_join(static_nest_w_res, by = c('well','round_depth','orientation')) %>%
  rename(upt_data = data.x, trx_data = data.y) %>%
  map2_df(.x = .$vs_model, .y = .$trx_data, 
          .f = ~augment(.x, newdata = .y, se_fit=TRUE)) %>%
  rename(pred_vs = .fitted, se_vs = .se.fit) %>%
  dplyr::select(pred_vs, se_vs)

merged_w_res <- cbind(vp_df, vs_df) %>%
  mutate(se_vp_per = se_vp/pred_vp) %>%
  mutate(se_vs_per = se_vs/pred_vs) %>%
  filter(pred_vp >0, pred_vs >0) %>%
  filter(!is.na(se_vp), !is.na(se_vs)) %>%
  filter(se_vp_per < 0.1) %>%
  filter(se_vs_per < 0.1)

write_csv(merged_w_res, '../data/merged_tests_w_residual.csv')
```
