---
title: "Geospatial Layer Preparation"
output: html_document
---

This study references numerous geospatial layers, including polygons,
points, and rasters. These layers are preprocessed here and saved in the data
folder for reproducibility and code efficiency. Below, vector-format 
layers (points, etc.) are saved as GeoJSON files and raster-format layers are 
saved as GeoTiffs. 

```{r setup, include=FALSE}
library(raster)
library(sf)
library(tidyverse)
library(janitor)
library(ggplot2)
library(dplyr)
```

## Setup study bounds

We setup a 100 km x 100 km grid for this study, with a resolution of 2.5 km x 2.5 km
(see the paper).

```{r}
study_bound_coords <- rbind(
  c(460000, 5980000), c(460000, 6080000), c(560000, 6080000), 
  c(560000, 5980000), c(460000, 5980000)
  )

# make a polygon
study_bounds <- list(study_bound_coords) %>%
  st_polygon() %>%
  st_sfc(., crs = 26911) %>%
  st_sf(.)

study_proj = st_crs(study_bounds)$proj4string

st_write(study_bounds, '../data/study_bounds.GeoJSON', delete_dsn = TRUE)
```

## Make a 2500x2500 m raster

We now transform the bounds into a grid using the raster package.

```{r}
study_raster <- raster(
  extent(study_bounds), 
  resolution = 2500,
  crs = st_crs(study_bounds)$proj4string
  )
study_raster[is.na(study_raster[])] <- 0 

writeRaster(study_raster, '../data/study_raster.tif', overwrite=TRUE)
```

## Creat a Duvernay-wide raster for mapping purposes

# Load surface locations

We load the surface hole locations from the AGS well list layer, filter to make
sure they are in the study bounds, and combine them with the AGS well list to
get both the license, wellname, uwi etc.

Original data can be found here: http://www1.aer.ca/ProductCatalogue/10.html

A copy is saved to data-raw which isn't uploaded to the repository. Generally, 
anything in `data-raw` is omitted due to licensing restrictions or politeness
of not reproducing an original system of record.

```{r}
surface_locs = sf::st_read(
  dsn = '../data-raw/sh_shapefile/', layer='ST37_SH_GCS_NAD83') %>% 
  st_transform(., crs = 26911)

study_locs = surface_locs %>% filter(
  st_within(surface_locs, study_bounds,sparse = FALSE)
  ) %>% janitor::clean_names()

well_list = read_tsv(
  '../data-raw/sh_shapefile/WellList.txt', 
  col_names = c('uwi_disp','uwi','flag','name','field','pool','os_area',
                'os_dep','licence', 'status', 'issue_data', 'code','agent', 
                'operator','drill_data', 'td_m','stat_code', 'stat_date', 
                'fluid', 'mode', 'struct', 'scheme', 'scheme_sub')
  )

well_list$uwi_clean = well_list$uwi_disp %>% 
  str_c(1,.) %>%
  str_replace(., "/", "") %>% 
  str_replace(., "/", "0") %>% 
  str_replace_all(., "-", "")

# Merge with surface locations
study_locs = study_locs %>% 
  merge(well_list %>% dplyr::select(licence, uwi = uwi_clean), 
        on='licence', how='left')

st_write(study_locs, '../data/study_surf_locs.GeoJSON', delete_dsn = TRUE)
```

Load the Duvernay isopach and write the study mask for predictions & figures:
http://www.ags.gov.ab.ca/publications/DIG/ZIP/DIG_2008_0124.zip

```{r}
dv_bounds <- st_read(dsn = '../data-raw/dvisopach/', layer = 'fg1217_py_ll') %>%
  st_transform(., crs = 26911) %>%
  janitor::clean_names() %>%
  dplyr::filter(descr1 != 'Dolomite') %>%
  dplyr::filter(descr1 != 'Mixed dolomite and evaporite') %>%
  dplyr::filter(descr1 != 'Limestone')

st_write(dv_bounds, '../data/entire_dv_bounds.GeoJSON', delete_dsn=TRUE)

study_dv_bounds <- dv_bounds %>%
  st_intersection(study_bounds)

st_write(study_dv_bounds, '../data/study_dv_bounds.GeoJSON', delete_dsn=TRUE)
```

Load the ground surface from Alberta Geological Survey's 3D Model
https://ags.aer.ca/publication/3d-pgf-model-v2

```{r}
ground_surface <- raster(
  '../data-raw/ags_geomodel/Continuous_Model_Horizons/cmh_01_QNgPg_sediment_top.asc'
  ) %>%
  raster::projectRaster(., crs = study_proj) %>%
  crop(., study_bounds)
names(ground_surface) <- 'ground_elev_masl'

# Verify projections visually
surface_df = ground_surface %>%
  rasterToPoints(., spatial=TRUE) %>%
  data.frame(.)

ggplot() +
  geom_tile(data=surface_df, aes(x=x,y=y,fill=ground_elev_masl)) +
  geom_sf(data=study_dv_bounds, fill=NA, colour='yellow')

writeRaster(ground_surface, '../data/ground_elev_masl.tif', overwrite=TRUE)
```

Load the Duvernay top from Alberta Geological Survey's 3D Model

```{r}
dv_top <- raster(
  '../data-raw/ags_geomodel/Discrete_Model_Horizons_TOP/dmh_54_D_Dv_Mu_top.asc'
  ) %>%
  raster::projectRaster(., crs = study_proj) %>%
  crop(., study_bounds)
names(dv_top) <- 'dv_top_masl'

# Verify projections visually
dv_top_df = dv_top %>%
  rasterToPoints(., spatial=TRUE) %>%
  data.frame(.)

ggplot() +
  geom_tile(data=dv_top_df, aes(x=x,y=y,fill=dv_top_masl)) +
  geom_sf(data=study_dv_bounds, fill=NA, colour='yellow')

writeRaster(dv_top, '../data/dv_top_masl.tif', overwrite=TRUE)
```
Load the Duvernay Base

```{r}
dv_base <- raster(
  '../data-raw/ags_geomodel/Discrete_Model_Horizons_BASE/dmh_54_D_Dv_Mu_base.asc'
  ) %>%
  raster::projectRaster(., crs = study_proj) %>%
  crop(., study_bounds)
names(dv_base) <- 'dv_base_masl'

# Verify projections visually
dv_base_df = dv_base %>%
  rasterToPoints(., spatial=TRUE) %>%
  data.frame(.)

ggplot() +
  geom_tile(data=dv_base_df, aes(x=x,y=y,fill=dv_base_masl)) +
  geom_sf(data=study_dv_bounds, fill=NA, colour='yellow')

writeRaster(dv_base, '../data/dv_base_masl.tif', overwrite=TRUE)
```

Get DV Isopach

```{r}
dv_isopach = dv_top - dv_base
names(dv_isopach) <- 'dv_isopach_m'

dv_isopach_df = dv_isopach %>%
  rasterToPoints(., spatial=TRUE) %>%
  data.frame(.)

ggplot() +
  geom_tile(data=dv_isopach_df, aes(x=x,y=y,fill=dv_isopach_m)) +
  geom_sf(data=study_dv_bounds, fill=NA, colour='yellow')

writeRaster(dv_isopach, '../data/dv_iso_m.tif', overwrite=TRUE)
```
