---
title: "Data Preparation"
output: html_document
---

```{r setup, include=FALSE}
library(duvernaygeomechmodel)
```

This notebook conducts exploratory data analysis on the experimental data

# Load files

```{r}
merged_no_res = read_csv('../data/merged_tests_no_residual.csv')
merged_w_res = read_csv('../data/merged_tests_w_residual.csv')

```

# Brazilian tests

Brazilian tests suggest that the carbonate facies has a higher tensile strength than the shale facies. But this is statistically rejected by a Welsh's t-test.

```{r}
brazilian = read_csv('../data/brazilian_tests.csv')

ggplot(brazilian) +
  geom_boxplot(aes(x = group, y = tensile_strength, colour = group)) +
  geom_jitter(aes(x = group, y = tensile_strength, colour = group), width = 0.25) +
  facet_grid(. ~ orientation) +
  labs(y = 'Tensile Strength (MPa)',
       x = '',
       title = '') +
  scale_color_manual(name = '', values = c('blue','dark green'))+
  ggsave('../output/brazilian_results.pdf', height = 6, width = 6) +
  ggsave('../output/brazilian_results.png', height = 6, width = 6)

brazilian %>%
  group_by(group) %>%
  summarize(n = n(),
            mean = mean(tensile_strength),
            sd = sd(tensile_strength))

t.test(tensile_strength ~ group, data = brazilian)
```

# UCS tests

The UCS results show significant strength anisotropy in both carbonate and shale formations.

```{r}
ucs = read_csv('../data/static_tests.csv') %>%
  filter(confining_pressure < 1)

ggplot(ucs) +
  geom_boxplot(aes(x = group, y = peak_stress_dev, colour = group)) +
  geom_jitter(aes(x = group, y = peak_stress_dev, colour = group), width = 0.25) +
  facet_grid(. ~ orientation) +
  labs(y = 'Unconfined Compressive Strength (MPa)',
       x = '',
       title = '') +
  scale_color_manual(name = '', values = c('blue','dark green'))+
  ggsave('../output/ucs_results.pdf', height = 6, width = 6) +
  ggsave('../output/ucs_results.png', height = 6, width = 6)

ucs %>%
  group_by(group, orientation) %>%
  summarize(n = n(),
            mean = mean(peak_stress_dev),
            sd = sd(peak_stress_dev))

# shale anisotropy test
t.test(
  peak_stress_dev ~ orientation, 
  data = ucs %>% filter(group == 'shale', orientation != 'oblique')
  )

# are oblique samples the weakest?
t.test(
  peak_stress_dev ~ orientation, 
  data = ucs %>% filter(group == 'shale', orientation != 'vertical')
  )

# carbonate anisotropy test
t.test(
  peak_stress_dev ~ orientation, 
  data = ucs %>% filter(group == 'carbonate', orientation != 'oblique')
  )

```

# Triaxial Results

```{r}
triaxial = read_csv('../data/static_tests.csv') %>%
  filter(confining_pressure > 0)

p1 = ggplot(triaxial) +
  geom_histogram(aes(x = youngs_modulus, fill = orientation, y = ..ncount..), bins = 50) +
  facet_grid(. ~ group) +
  coord_flip() +
  labs(y = 'Normalized Count',
       x = 'Young\'s Modulus (GPa)',
       title = '(a) Young\'s Modulus') +
  scale_color_manual(name = '', values = c('blue','dark green')) +
  theme(legend.position = "bottom")

p2 = ggplot(triaxial) +
  geom_histogram(aes(x = poisson_ratio, fill = orientation, y = ..ncount..), bins = 50) +
  facet_grid(. ~ group) +
  coord_flip() +
  labs(y = 'Normalized Count',
       x = 'Poisson\'s Ratio',
       title = '(b) Poisson\'s Ratio') +
  scale_color_manual(name = '', values = c('blue','dark green')) +
  theme(legend.position = "bottom")

p3 = ggplot(triaxial) +
  geom_boxplot(aes(x = group, y = bulk_density, colour = group)) +
  geom_jitter(aes(x = group, y = bulk_density, colour = group), width = 0.25) +
  ylim(2.4,2.8) +
  labs(y = 'Bulk Density (g/cc)',
       x = '',
       title = '(c) Bulk Density') +
  scale_color_manual(name = '', values = c('blue','dark green')) +
  theme(legend.position = "bottom")

grobs = arrangeGrob(grobs=mat, widths = unit(c(4.5,4.5,3),"in"), heights = unit(c(6),"in"))

ggsave("../output/triaxial_results.pdf", grobs, width = 12, height = 6)
ggsave("../output/triaxial_results.png", grobs, width = 12, height = 6, dpi=600)
```
