---
title: "Geostatistics CoSimulation"
author: "Scott McKean"
date: "05/02/2021"
output: html_document
---

# Stress Cosimulation

This document goes over development of a LMC and cosimulation of the logs, 
conditioned on the vertical stress exhaustive data

```{r setup, include=FALSE}
# geospatial packages
library(sf)
library(gstat)
library(raster)
library(sp)

# statistical normalization and analysis
library(ggpubr)
library(moments)
library(bestNormalize)

# data munging & plotting
library(tidyverse)
library(viridis)

source('../R/geospatial.R')
```

## Load spatial dataframes

See 'geostats_data_prep.Rmd' to generate these files and dump into the cache.
Read GeoJSONs from the cache.

We run an aggregation function to take the mean gradient across measurement
points because a) we aren't doing 3D kriging, and b) because geostatistics doesn't
work with zero distance variance that isn't equal to zero

```{r cars}
study_raster = raster('../data/study_raster.tif')
sv_sf = st_read('../data/sv_sf.GeoJSON')
logs_sf = st_read('../data/facies_data.GeoJSON')
shale_sf = logs_sf %>% filter(facies == 'dv_shale')
shale_sp = as_Spatial(shale_sf)
carb_sf = logs_sf %>% filter(facies == 'dv_carb')
carb_sp = as_Spatial(carb_sf)

```
# Variography

This section investigates the normality, anisotropy, and variography of each of 
the log variables (bulk density, vp, vs). We also repeat the SV variography
minus the exploration (see geostats_stress.Rmd)

Since Sequential Gaussian Cosimulation is used, we transform variables to normal 
distributions using the ordered quantile technique 

## Descriptive Variography - SV

```{r}
sv_sp = as_Spatial(sv_sf)
sv_sp = sv_sp[!is.na(sv_sp$svgrad_kpa_m),]
sv_norm = bestNormalize::orderNorm(sv_sp$svgrad_kpa_m)
sv_sp$svgrad_norm = sv_sp$svgrad_norm = sv_norm$x.t
sv_v = variogram(svgrad_norm~1, sv_sp)
sv_vgm = vgm(0.69, "Sph", 50E3, nugget = 0.01)
sv_vgm = vgm(0.3, "Sph", 15E3, anis=c(67, 0.75), add.to=sv_vgm)
plot(sv_v, sv_vgm)
```
## Descriptive Variography - RHOB / Shale

- Density actually has a relatively well behaved variogram
- No anisotropy noted
- Good coverage
- Nothing of note in variogram map

```{r}
shale_rhob = shale_sp[!is.na(shale_sp$rhob_mean),]
# investigate normality

# make anisotropic variograms
ggdensity(shale_sf, x='rhob_mean') +
  stat_overlay_normal_density(color='red', linetype='dashed')
moments::skewness(shale_sf$rhob_mean)
moments::kurtosis(shale_sf$rhob_mean)

# normal transform
rhob_norm = bestNormalize::orderNorm(shale_rhob$rhob_mean)
shale_rhob$rhob_norm = rhob_norm$x.t
ggdensity(as.data.frame(shale_rhob), x='rhob_norm') +
  stat_overlay_normal_density(color='red', linetype='dashed')
moments::skewness(rhob_norm$x.t)
moments::kurtosis(rhob_norm$x.t)

# what is the global variance (should be 1...)
var(shale_rhob$rhob_norm)

# make an experimental variogram
shale_rhob_v = variogram(rhob_norm~1, shale_rhob)
plot(shale_rhob_v)

# check directional variograms
rhob_v_dir = variogram(rhob_norm~1, shale_rhob, alpha=c(0,22,45,67,90,112,135,157))
plot(rhob_v_dir)

# check for geometric anisotropy
ggvariogram_map(
  formula = rhob_norm~1, 
  sp_df = shale_rhob,
  cutoff = 100E3, 
  width = 5E3,
  threshold = 1
  )

shale_rhob_vgm = vgm(0.8, "Sph", 16E3, nugget = 0.2)
plot(shale_rhob_v, shale_rhob_vgm)
```

## Descriptive Variography - RHOB / Carb

- The carbonate density appears to be almost all nugget (60% of variance?)
- Some short range predictability is noted

```{r}
carb_rhob = carb_sp[!is.na(carb_sp$rhob_mean),]
# investigate normality

# make anisotropic variograms
ggdensity(carb_sf, x='rhob_mean') +
  stat_overlay_normal_density(color='red', linetype='dashed')
moments::skewness(carb_sf$rhob_mean)
moments::kurtosis(carb_sf$rhob_mean)

# normal transform
rhob_norm = bestNormalize::orderNorm(carb_sp$rhob_mean)
carb_sf$rhob_norm = carb_sp$rhob_norm = rhob_norm$x.t
ggdensity(carb_sf, x='rhob_norm') +
  stat_overlay_normal_density(color='red', linetype='dashed')
moments::skewness(rhob_norm$x.t)
moments::kurtosis(rhob_norm$x.t)

# what is the global variance (should be 1...)
var(carb_sp$rhob_norm)

# make an experimental variogram
carb_rhob_v = variogram(rhob_norm~1, carb_sp)
plot(carb_rhob_v)

# check directional variograms
rhob_v_dir = variogram(rhob_norm~1, carb_sp, alpha=c(0,22,45,67,90,112,135,157))
plot(rhob_v_dir)

# check for geometric anisotropy
ggvariogram_map(
  formula = rhob_norm~1, 
  sp_df = carb_sp,
  cutoff = 100E3, 
  width = 5E3,
  threshold = 1
  )

carb_rhob_vgm = vgm(0.4, "Sph", 16E3, nugget = 0.6)
plot(carb_rhob_v, carb_rhob_vgm)
```
## Descriptive Variography - DTC / Shale

- Density actually has a relatively well behaved variogram
- No anisotropy noted
- Good coverage
- Nothing of note in variogram map

```{r}
shale_dtc = shale_sp[!is.na(shale_sp$dtc_mean),]
# investigate normality

# make anisotropic variograms
ggdensity(shale_sf, x='dtc_mean') +
  stat_overlay_normal_density(color='red', linetype='dashed')
moments::skewness(shale_dtc$dtc_mean)
moments::kurtosis(shale_dtc$dtc_mean)

# normal transform
dtc_norm = bestNormalize::orderNorm(shale_dtc$dtc_mean)
shale_dtc$dtc_norm = dtc_norm$x.t
ggdensity(as.data.frame(shale_dtc), x='dtc_norm') +
  stat_overlay_normal_density(color='red', linetype='dashed')
moments::skewness(rhob_norm$x.t)
moments::kurtosis(rhob_norm$x.t)

# what is the global variance (should be 1...)
var(shale_dtc$dtc_norm)

# make an experimental variogram
shale_dtc_v = variogram(dtc_norm~1, shale_dtc)
plot(shale_dtc_v)

# check directional variograms
shale_dtc_v_dir = variogram(dtc_norm~1, shale_dtc, alpha=c(0,22,45,67,90,112,135,157))
plot(shale_dtc_v_dir)

# check for geometric anisotropy
ggvariogram_map(
  formula = dtc_norm~1, 
  sp_df = shale_dtc,
  cutoff = 100E3, 
  width = 5E3,
  threshold = 1
  )

shale_dtc_vgm = vgm(0.6, "Sph", 40E3, nugget = 0.4)
plot(shale_dtc_v, shale_dtc_vgm)
```

## Descriptive Variography - DTC / Carb

- The carbonate DTC appears shorter range than shale
- Some anisotropy is noted, but nothing enough to be significant or justify
the use of anisotropic variograms
- The range of 16 km from Carb/Rhob matches really nicely with the DTC.

```{r}
carb_dtc = carb_sp[!is.na(carb_sp$dtc_mean),]
# investigate normality

# make anisotropic variograms
ggdensity(carb_sf, x='dtc_mean') +
  stat_overlay_normal_density(color='red', linetype='dashed')
moments::skewness(carb_dtc$dtc_mean)
moments::kurtosis(carb_dtc$dtc_mean)

# normal transform
dtc_norm = bestNormalize::orderNorm(carb_dtc$dtc_mean)
carb_dtc$dtc_norm = dtc_norm$x.t
ggdensity(as.data.frame(carb_dtc), x='dtc_norm') +
  stat_overlay_normal_density(color='red', linetype='dashed')
moments::skewness(dtc_norm$x.t)
moments::kurtosis(dtc_norm$x.t)

# what is the global variance (should be 1...)
var(carb_dtc$dtc_norm)

# make an experimental variogram
carb_dtc_v = variogram(dtc_norm~1, carb_dtc)
plot(carb_dtc_v)

# check directional variograms
dtc_v_dir = variogram(dtc_norm~1, carb_dtc, alpha=c(0,22,45,67,90,112,135,157))
plot(dtc_v_dir)

# check for geometric anisotropy
ggvariogram_map(
  formula = dtc_norm~1, 
  sp_df = carb_dtc,
  cutoff = 100E3, 
  width = 5E3,
  threshold = 1
  )

carb_dtc_vgm = vgm(0.65, "Sph", 16E3, nugget = 0.35)
plot(carb_dtc_v, carb_dtc_vgm)
```

## Descriptive Variography - DTS / Shale

- DTS shale is pretty damn normal
- We only have 24 observations here... low numbers, result in some strange
variograms
- We use the same scale as used on the DTC shale (40km), with a decent nugget

```{r}
shale_dts = shale_sp[!is.na(shale_sp$dts_mean),]
# investigate normality

# make anisotropic variograms
ggdensity(shale_sf, x='dts_mean') +
  stat_overlay_normal_density(color='red', linetype='dashed')
moments::skewness(shale_dts$dts_mean)
moments::kurtosis(shale_dts$dts_mean)

# normal transform
dts_norm = bestNormalize::orderNorm(shale_dts$dts_mean)
shale_dts$dts_norm = dts_norm$x.t
ggdensity(as.data.frame(shale_dts), x='dts_norm') +
  stat_overlay_normal_density(color='red', linetype='dashed')
moments::skewness(rhob_norm$x.t)
moments::kurtosis(rhob_norm$x.t)

# what is the global variance (should be 1...)
var(shale_dts$dts_norm)

# make an experimental variogram
shale_dts_v = variogram(dts_norm~1, shale_dts)
plot(shale_dts_v)

# check directional variograms
shale_dts_v_dir = variogram(dts_norm~1, shale_dts, alpha=c(0,22,45,67,90,112,135,157))
plot(shale_dts_v_dir)

# check for geometric anisotropy
ggvariogram_map(
  formula = dts_norm~1, 
  sp_df = shale_dts,
  cutoff = 100E3, 
  width = 5E3,
  threshold = 1
  )

shale_dts_vgm = vgm(0.75, "Sph", 40E3, nugget = 0.25)
plot(shale_dts_v, shale_dts_vgm)
```

## Descriptive Variography - dts / Carb

- Carbonate DTS is skewed towards the lower end of the scale
- We only have 17 observations...
- Once again, we see a smaller range (16 km should do fine)
- With very limited information, we start with the same variogram as we
use for DTC/Carb.

```{r}
carb_dts = carb_sp[!is.na(carb_sp$dts_mean),]
# investigate normality

# make anisotropic variograms
ggdensity(carb_sf, x='dts_mean') +
  stat_overlay_normal_density(color='red', linetype='dashed')
moments::skewness(carb_dts$dts_mean)
moments::kurtosis(carb_dts$dts_mean)

# normal transform
dts_norm = bestNormalize::orderNorm(carb_dts$dts_mean)
carb_dts$dts_norm = dts_norm$x.t
ggdensity(as.data.frame(carb_dts), x='dts_norm') +
  stat_overlay_normal_density(color='red', linetype='dashed')
moments::skewness(dts_norm$x.t)
moments::kurtosis(dts_norm$x.t)

# what is the global variance (should be 1...)
var(carb_dts$dts_norm)

# make an experimental variogram
carb_dts_v = variogram(dts_norm~1, carb_dts)
plot(carb_dts_v)

# check directional variograms
dts_v_dir = variogram(dts_norm~1, carb_dts, alpha=c(0,22,45,67,90,112,135,157))
plot(dts_v_dir)

# check for geometric anisotropy
ggvariogram_map(
  formula = dts_norm~1, 
  sp_df = carb_dts,
  cutoff = 100E3, 
  width = 5E3,
  threshold = 1
  )

carb_dts_vgm = vgm(0.65, "Sph", 16E3, nugget = 0.35)
plot(carb_dts_v, carb_dts_vgm)
```
## Cosimulation 

We are going to cosimulate stress values using a Linear Model of Coregionalization (LMC)
via gstat. We could get fancier, but that's not what is important, what is important
is that we have heterotopic data underpinned by an exhaustive secondary variable
and need to quantify the uncertainty in our stress measurements, and how that 
compares to our model uncertainty.

We first try with two variables to develop intuition about cross-variogram fitting

```{r}
# make a prediction grid
st_grid <- rasterToPoints(study_raster, spatial = TRUE)
gridded(st_grid) <- TRUE
st_grid <- as(st_grid, "SpatialPixels")
```

For the logs we have

SV, RHOB, DTC, DTS
SV is exhaustive and the same for both carbonate and shale (the exhaustive link raster)
RHOB, DTC, DTS are simulated separately for carbonate and shale

Shale first 
```{r}
# secondary (SV)
sv_vgm = vgm(0.3, "Sph", 50E3, 0.2)
sv_vgm = vgm(0.5, "Lin", 10E3, add.to=sv_vgm)
g = gstat(id='sv', formula=svgrad_norm~1, data=sv_sp, beta=0, nmax=50)
g = gstat(g, id='sv', model=sv_vgm)

rhob_vgm = vgm(0.3, "Sph", 50E3, 0.2)
rhob_vgm = vgm(0.5, "Lin", 10E3, add.to=rhob_vgm)
g = gstat(g, id='rhob', formula=rhob_norm~1, data=shale_rhob, beta=0, nmax=50)
g = gstat(g, id='rhob', model=rhob_vgm)

dtc_vgm = vgm(0.5, "Sph", 50E3, 0.2)
dtc_vgm = vgm(0.5, "Lin", 10E3, add.to=dtc_vgm)
g = gstat(g, id='dtc', formula = dtc_norm~1, data=shale_dtc, beta=0, nmax=50)
g = gstat(g, id='dtc', model=dtc_vgm)

dts_vgm = vgm(0.5, "Sph", 50E3, 0.2)
dts_vgm = vgm(0.5, "Lin", 10E3, add.to=dts_vgm)
g = gstat(g, id='dts', formula = dts_norm~1, data=shale_dts, beta=0, nmax=50)
g = gstat(g, id='dts', model=dts_vgm)

cross_vgm = vgm(0.5, "Sph", 50E3, 0.2)
cross_vgm = vgm(0.5, "Lin", 10E3, add.to=cross_vgm)
g = gstat(g, id=c('sv','rhob'), model=cross_vgm, beta=0, nmax=50)
g = gstat(g, id=c('sv','dtc'), model=cross_vgm, beta=0, nmax=50)
g = gstat(g, id=c('sv','dts'), model=cross_vgm, beta=0, nmax=50)
g = gstat(g, id=c('rhob','dtc'), model=cross_vgm, beta=0, nmax=50)
g = gstat(g, id=c('rhob','dts'), model=cross_vgm, beta=0, nmax=50)
g = gstat(g, id=c('dtc','dts'), model=cross_vgm, beta=0, nmax=50)

v = variogram(g)

g = fit.lmc(v,g, fit.lmc =TRUE, fit.method=1)

plot(v,g)
g
```

Do conditional Gaussian cosimulation with three variables, a local neighbourhood of 50 points.

The LMC is likely a poor representation, representing a maximum likelihood estimation given positive definiteness, however, a Bayesian approach would likely be better... but for the purposes of this study we don't care. Honestly, the default settings aren't important because it is so difficult to fit a positive definite matrix with the variograms we see...

This fit isn't significantly improved by simulating two variables at a time, so we stick with a full cosimulation with LMC, despite a poor fit of the LMC. Bayesian methods may be able to improve upon this, but it still serves the point of quantifying uncertainty.

- Matrices are positive definite
- Matrices pass Cauchy-Schwartz inequality
- Simulation relies primarily on the cross-variograms, which means the absolute values may very but the relative difference should be minimized, which is very important for this study.

```{r}
g$set = list(nocheck=1)
pred = predict(g, st_grid, nsim=10)
```
