---
title: "Geostatistics Data Preparation"
author: "Scott McKean"
date: "05/02/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
library(sf)
library(tidyverse)
library(janitor)
library(plot3D)
library(viridis)
library(raster)
```

Setup our study bounds and make a 2.5 km x 2.5 km raster grid

```{r}
# set study bounds
study_bound_coords <- rbind(
  c(460000, 5980000), c(460000, 6080000), c(560000, 6080000), 
  c(560000, 5980000), c(460000, 5980000)
  )

# make a polygon
study_bounds <- list(study_bound_coords) %>%
  st_polygon() %>%
  st_sfc(., crs = 26911) %>%
  st_sf(.)

study_proj = st_crs(study_bounds)$proj4string

# make a 2500x2500 m raster
study_raster <- raster(
  extent(study_bounds), 
  resolution = 2500,
  crs = st_crs(study_bounds)$proj4string
  )
```

Load the Duvernay isopach
http://www.ags.gov.ab.ca/publications/DIG/ZIP/DIG_2008_0124.zip

```{r}
duvernay_iso <- st_read(dsn = '../data/dvisopach/', layer = 'fg1217_py_ll') %>%
  st_transform(., crs = 26911) %>%
  janitor::clean_names() %>%
  dplyr::filter(descr1 != 'Dolomite') %>%
  st_intersection(study_bounds)
```

Load the ground surface from Alberta Geological Survey's 3D Model
https://ags.aer.ca/publication/3d-pgf-model-v2

```{r}
ground_surface <- raster('../data/ags3dmodel/cmh_01_QNgPg_sediment_top.asc') %>%
  raster::projectRaster(., crs = study_proj)
```

Load Shen et al. (2018)'s  vertical stress data. Because the datum is provided in masl,
we need depth below ground surface to calculate a proper gradient. We use the raster
package to extract this from AGS's 3D geomodel using 

```{r}
sv_sf <- read_tsv('../data/shen2018/DIG_2018_0013_vertical_stress.txt') %>%
  st_as_sf(.,coords = c('Long_NAD83','Lat_NAD83'), crs = 4269) %>%
  st_transform(crs = 26911) %>%
  st_crop(study_bounds) %>%
  janitor::clean_names() %>%
  rename(datum_elev_masl = datum_el) %>%
  dplyr::mutate(ground_elev_masl = raster::extract(ground_surface, .,)) %>%
  dplyr::mutate(datum_mbgs = ground_elev_masl - datum_elev_masl) %>%
  mutate(sv_mpa = vertical_stress/1000) %>%
  mutate(svgrad_kpa_m = vertical_stress/datum_mbgs) %>%
  dplyr::select(datum_elev_masl, datum_mbgs, sv_mpa, svgrad_kpa_m)

st_write(sv_sf, "../cache/sv_sf.GeoJSON",delete_dsn = TRUE)
```

Show 2D and 3D gradients using viridis scale. The 3D plot demonstrates vertical variability across study area and shows why stress gradient is so important.

```{r}
ggplot(sv_sf) +
  geom_sf(data = duvernay_iso) +
  geom_sf(data = study_bounds, fill = NA) +
  geom_sf(aes(colour=svgrad_kpa_m)) +
  theme_minimal() +
  scale_colour_viridis(option='viridis') +
  coord_sf(datum=st_crs(26911)) +
  ggsave('../figure_outputs/svgrad_2d.pdf')

make_3d_plot(
  sv_sf, 
  z_col='datum_elev_masl', 
  col_col = 'svgrad_kpa_m',
  col_func = viridis(50),
  size=0.35
  )
```

Load Shen et al. (2018)'s shmin data. We assume a kelly bushing height of 7 m, 
which should be pretty fair.

```{r}
shmin_sf <- read_tsv('../data/shen2018/DIG_2018_0013_DFITSh.txt') %>%
  st_as_sf(.,coords = c('Long_NAD83','Lat_NAD83'), crs = 4269) %>%
  st_transform(crs= 26911) %>%
  dplyr::filter(st_contains(study_bounds, ., sparse = FALSE)) %>%
  janitor::clean_names() %>%
  dplyr::mutate(shmin = ifelse(
    is.na(sh_reported), 
    (p_vs_log_t_high + p_vs_log_t_low)/2,
    sh_reported)) %>%
  dplyr::mutate(datum_elev_masl = kb_el_masl - datum_tvd) %>%
  dplyr::mutate(datum_mbgs = datum_tvd - 7) %>%
  dplyr::mutate(uwi = str_remove_all(well_id,"-")) %>%
  dplyr::mutate(uwi = str_remove_all(uwi,"/")) %>%
  dplyr::mutate(shmin_mpa = shmin/1000) %>%
  dplyr::mutate(shmingrad_kpa_m = shmin / (datum_tvd - 7)) %>%
  dplyr::select(datum_elev_masl, datum_mbgs, shmin_mpa, shmingrad_kpa_m)

st_write(shmin_sf, "../cache/shmin_sf.GeoJSON", delete_dsn = TRUE)
```

Load Shen's 2018 pore pressure data. Following their approach, we take static
gradients when available, followed by the mean of the buildup analysis methods,
followed by the initial reported pore pressure.

```{r}
pp_sf <- read_tsv('../data/shen2018/DIG_2018_0013_pore_pressure.txt') %>%
  st_as_sf(.,coords = c('Long_NAD83','Lat_NAD83'), crs = 4269) %>%
  st_transform(crs= 26911) %>%
  dplyr::filter(st_contains(study_bounds, ., sparse = FALSE)) %>%
  janitor::clean_names() %>%
  dplyr::mutate(pp = case_when(
    !is.na(pore_static_gradient) ~ pore_static_gradient,
    test_type != 'static gradient' ~ mean(
      c(buildup_linear,buildup_radial,buildup_bi_linear),na.rm=TRUE),
    TRUE ~ reported_pore_initial,
  )) %>%
  dplyr::mutate(datum_elev_masl = kb_el_masl - datum_tvd) %>%
  dplyr::mutate(datum_mbgs = datum_tvd - 7) %>%
  dplyr::mutate(uwi = str_remove_all(well_id,"-")) %>%
  dplyr::mutate(uwi = str_remove_all(uwi,"/")) %>%
  dplyr::mutate(pp_mpa = pp/1000) %>%
  dplyr::mutate(ppgrad_kpa_m = pp/(datum_tvd-7)) %>%
  dplyr::select(datum_elev_masl, datum_mbgs, pp_mpa, ppgrad_kpa_m)

st_write(pp_sf, "../cache/pp_sf.GeoJSON", delete_dsn=TRUE)
```

Show 2D and 3D gradients using magma scale. Really shows the sparsity of the 
stress data. We overlay the pp/shmin on the other plot for reference to show 
the observation heterotopy.

```{r}
ggplot(shmin_sf) +
  geom_sf(data = duvernay_iso) +
  geom_sf(data = study_bounds, fill = NA) +
  geom_sf(data=pp_sf, colour='black', shape=21, size=3) +
  geom_sf(aes(colour=shmingrad_kpa_m),size=1.5) +
  theme_minimal() +
  scale_colour_viridis(option='magma') +
  coord_sf(datum=st_crs(26911)) +
  ggsave('../figure_outputs/shmingrad_2d.pdf')

make_3d_plot(
  shmin_sf, 
  z_col='datum_elev_masl', 
  col_col = 'shmingrad_kpa_m',
  col_func = magma(50),
  size=1
  )
```

Show 2D and 3D gradients using magma scale. Really shows the sparsity of the 
pore pressure data.

```{r}
ggplot(pp_sf) +
  geom_sf(data = duvernay_iso) +
  geom_sf(data = study_bounds, fill = NA) +
  geom_sf(data=shmin_sf, colour='black', shape=21, size=3) +
  geom_sf(aes(colour=ppgrad_kpa_m), size=1.5) +
  theme_minimal() +
  scale_colour_viridis(option='plasma') +
  coord_sf(datum=st_crs(26911)) +
  ggsave('../figure_outputs/ppgrad_kpa_m.pdf')

make_3d_plot(
  pp_sf, 
  z_col='datum_elev_masl', 
  col_col = 'ppgrad_kpa_m',
  col_func = plasma(50),
  size=1
  )
```

Now that we have the stress data, save as R objects so we can load into the
geostatistics script.

```{r}

load('pp_sf.rds')
rm(pp_sf)
```
