---
title: "Geostatistics Data Preparation"
author: "Scott McKean"
date: "05/02/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

This notebook loads the raw stress data from Shen et al. (2018)'s publication, 
produced some figures for the paper, and outputs GeoJSON files for the geostatistics
workflow.

```{r setup, include=FALSE}
library(duvernaygeomechmodel)
library(sf)
library(tidyverse)
library(janitor)
library(plot3D)
library(viridis)
library(raster)

# From geospatial_layers_prep.Rmd
study_bounds = st_read('../data/study_bounds.GeoJSON')
study_proj = st_crs(study_bounds)$proj4string
study_raster = raster::raster('../data/study_raster.tif')
study_dv_bounds = st_read('../data/study_dv_bounds.GeoJSON')

ground_surface = raster::raster('../data/ground_elev_masl.tif')
```

Load vertical stress data. Because the datum is provided in masl,
we need depth below ground surface to calculate a proper gradient. We use the raster
package to extract this from AGS's 3D geomodel using 

```{r}
sv_sf <- read_tsv('../data-raw/shen2018/DIG_2018_0013_vertical_stress.txt') %>%
  st_as_sf(.,coords = c('Long_NAD83','Lat_NAD83'), crs = 4269) %>%
  st_transform(crs = 26911) %>%
  st_crop(study_bounds) %>%
  janitor::clean_names() %>%
  rename(datum_elev_masl = datum_el) %>%
  dplyr::mutate(ground_elev_masl = raster::extract(ground_surface, .,)) %>%
  dplyr::mutate(datum_mbgs = ground_elev_masl - datum_elev_masl) %>%
  mutate(sv_mpa = vertical_stress/1000) %>%
  mutate(svgrad_kpa_m = vertical_stress/datum_mbgs) %>%
  dplyr::select(datum_elev_masl, datum_mbgs, sv_mpa, svgrad_kpa_m)

st_write(sv_sf, "../data/sv_sf.GeoJSON", delete_dsn=TRUE)
```

Show 2D and 3D gradients using viridis scale. The 3D plot demonstrates vertical 
variability across study area and shows why stress gradient is so important.

```{r}
ggplot(sv_sf) +
  geom_sf(data = study_dv_bounds) +
  geom_sf(data = study_bounds, fill = NA) +
  geom_sf(aes(colour=svgrad_kpa_m)) +
  theme_minimal() +
  scale_colour_viridis(option='viridis') +
  coord_sf(datum=st_crs(26911)) +
  ggsave('../figure_outputs/svgrad_2d.pdf')

make_3d_plot(
  sv_sf, 
  z_col='datum_elev_masl', 
  col_col = 'svgrad_kpa_m',
  col_func = viridis(50),
  size=0.35
  )
```

Load Shen et al. (2018)'s shmin data. We assume a kelly bushing height of 7 m, 
which should be pretty fair.

```{r}
shmin_sf <- read_tsv('../data-raw/shen2018/DIG_2018_0013_DFITSh.txt') %>%
  st_as_sf(.,coords = c('Long_NAD83','Lat_NAD83'), crs = 4269) %>%
  st_transform(crs= 26911) %>%
  dplyr::filter(st_contains(study_bounds, ., sparse = FALSE)) %>%
  janitor::clean_names() %>%
  dplyr::mutate(shmin = ifelse(
    is.na(sh_reported), 
    (p_vs_log_t_high + p_vs_log_t_low)/2,
    sh_reported)) %>%
  dplyr::mutate(datum_elev_masl = kb_el_masl - datum_tvd) %>%
  dplyr::mutate(datum_mbgs = datum_tvd - 7) %>%
  dplyr::mutate(uwi = str_remove_all(well_id,"-")) %>%
  dplyr::mutate(uwi = str_remove_all(uwi,"/")) %>%
  dplyr::mutate(shmin_mpa = shmin/1000) %>%
  dplyr::mutate(shmingrad_kpa_m = shmin / (datum_tvd - 7)) %>%
  dplyr::select(datum_elev_masl, datum_mbgs, shmin_mpa, shmingrad_kpa_m)

st_write(shmin_sf, "../data/shmin_sf.GeoJSON", delete_dsn = TRUE)
```

Load Shen's 2018 pore pressure data. Following their approach, we take static
gradients when available, followed by the mean of the buildup analysis methods,
followed by the initial reported pore pressure.

```{r}
pp_sf <- read_tsv('../data-raw/shen2018/DIG_2018_0013_pore_pressure.txt') %>%
  st_as_sf(.,coords = c('Long_NAD83','Lat_NAD83'), crs = 4269) %>%
  st_transform(crs= 26911) %>%
  dplyr::filter(st_contains(study_bounds, ., sparse = FALSE)) %>%
  janitor::clean_names() %>%
  dplyr::mutate(pp = case_when(
    !is.na(pore_static_gradient) ~ pore_static_gradient,
    test_type != 'static gradient' ~ mean(
      c(buildup_linear,buildup_radial,buildup_bi_linear),na.rm=TRUE),
    TRUE ~ reported_pore_initial,
  )) %>%
  dplyr::mutate(datum_elev_masl = kb_el_masl - datum_tvd) %>%
  dplyr::mutate(datum_mbgs = datum_tvd - 7) %>%
  dplyr::mutate(uwi = str_remove_all(well_id,"-")) %>%
  dplyr::mutate(uwi = str_remove_all(uwi,"/")) %>%
  dplyr::mutate(pp_mpa = pp/1000) %>%
  dplyr::mutate(ppgrad_kpa_m = pp/(datum_tvd-7)) %>%
  dplyr::select(datum_elev_masl, datum_mbgs, pp_mpa, ppgrad_kpa_m)

st_write(pp_sf, "../cache/pp_sf.GeoJSON", delete_dsn=TRUE)
```

Show 2D and 3D gradients using magma scale. Really shows the sparsity of the 
stress data. We overlay the pp/shmin on the other plot for reference to show 
the observation heterotopy.

```{r}
ggplot(shmin_sf) +
  geom_sf(data = study_dv_bounds) +
  geom_sf(data = study_bounds, fill = NA) +
  geom_sf(data=pp_sf, colour='black', shape=21, size=3) +
  geom_sf(aes(colour=shmingrad_kpa_m),size=1.5) +
  theme_minimal() +
  scale_colour_viridis(option='magma') +
  coord_sf(datum=st_crs(26911)) +
  ggsave('../figure_outputs/shmingrad_2d.pdf')

make_3d_plot(
  shmin_sf, 
  z_col='datum_elev_masl', 
  col_col = 'shmingrad_kpa_m',
  col_func = magma(50),
  size=1
  )
```

Show 2D and 3D gradients using magma scale. Really shows the sparsity of the 
pore pressure data.

```{r}
ggplot(pp_sf) +
  geom_sf(data = study_dv_bounds) +
  geom_sf(data = study_bounds, fill = NA) +
  geom_sf(data=shmin_sf, colour='black', shape=21, size=3) +
  geom_sf(aes(colour=ppgrad_kpa_m), size=1.5) +
  theme_minimal() +
  scale_colour_viridis(option='plasma') +
  coord_sf(datum=st_crs(26911)) +
  ggsave('../figure_outputs/ppgrad_kpa_m.pdf')

make_3d_plot(
  pp_sf, 
  z_col='datum_elev_masl', 
  col_col = 'ppgrad_kpa_m',
  col_func = plasma(50),
  size=1
  )
```
